Evolution of Income Inequality in Canada

### 2022-04-28 outputs/paper.Rmd
fix grammar

In recent decades the world as a whole is developing fast economically, and one evidence of which is the rise in income level. As more countries become more developed in terms of their economy and eventually overcome their issues of poverty, their governments have gradually shift their goal to dealing with the problem of economic inequality. 
It concerns how the wealth and income of the population is distributed. Although we are generally living higher quality lives and getting higher incomes, another problem is being put forward by people who are claiming that the gap between the rich and poor is getting larger due to technological advancement, globalization, countries' governmental structure and policies, which betters off the people who are already wealthy by making it easy for them to seek for more opportunities, thus, snowballing their assets faster. Income disparity can be a huge issue. It makes the income per capita of a country less reflective of its people's true living quality, because income per capita does not tell the distribution of the income. The richest people may pull up the statistics about the average/median income level, but it doesn't mean the country is performing well economically as a whole. In addition to that, income inequality also have other potential negative effects. It could reduce the stability of the economy and increase the risk of financial crisis. It could also result in corruption, mis-allocation of resources as the rich is empowered economically, socially and even politically [@cite1].
Canada is one of the country in the world that provides the most benefits and welfare, yet people still claimed that the economic inequality is rising along with the income over the past 2 decades [@cite2]. Therefore, I'm interested in whether income inequality is a serious issue in Canada. If so, in which regions and what period of time is it a serious issue? And is it possible to predict how the disparity of income change in future?

The data used in this report is obtained from Open Government Data of Statistics Canada. The dataset is called "Upper income limit, income share and average income by economic family type and income decile", and it provides different kinds information about the income level of the households of Canada between 1976 and 2020 by the Canadian Income Survey. In this dataset, I'm interested in the year the data is from, the geographical location, the income decile so that we could observe the difference between the lowest and highest decile, and the actual value of the income. In section \@ref(data), a cleaned dataset is obtained and explained to perform further analysis. Visualizations in the forms of tables and figures are presented to help to explain the possible correlation between year, provinces and the income disparity in Canada. By data visualization, I've primarily assumed that the income disparity in Canada has generally increased as time passed and presents different trends in different provinces before constructing the model. In section \@ref(model), a multiple linear regression model is constructed to justify the relationship between year, province and income inequality, and to make future predictions. The interpretation of the final model along with all the findings regarding the evolution of income inequality is presented in section \@ref(results). A discussion is carried out in section \@ref(discussion) on the implications of the findings regarding time and geographical location, and possible solutions to this issue in federal and provincial level, as well as the weaknesses and future steps of this paper.

@@

The dataset, which is a summary of the income-related statistics in Canada between 1976 and 2020, is obtained from the Open Government Portal of Statistics Canada. The source of the data is from the Canadian Income Survey conducted annually. The survey covers all the possible population in Canada except for the people and households living in remote areas or indigenous settlements. It doesn't really affect the outcome of the survey, because these people only take up less than 2 percent of the population [@cite3]. The selected respondents were drawn from the Labour Force Survey samples, which is based on a stratified probability sampling [@cite3]. To reduce some non-sampling errors, telephone interviews ware conducted prior to the main survey to increase the response rate of the selected participants Once they give consent to completing the survey, they will be able to complete it in an online form. Respondents are protected from the confidentiality rules, so that their privacy won't be disclosed after the survey is made public.
One possible bias came with the survey methodology was the over-coverage of the units that are not the target population and the under-coverage of certain sub-population that should be included in the survey, such as certain remote areas.

@@

In order to get further familiarized with the dataset and estimate the possible associations between the nation's average income and other factors, exploratory analysis is carried out by conducting data visualizations to observe whether the patterns between certain factors matches the generally expectations of the trend of income and income inequality in Canada. First of all, we will have a look through the overall trend of income level in Canada.

@@

As Figure \@ref(fig:figure1) has demonstrated, Canadians have experienced a rapidly growing income level since 1976 to now. Early 1980s and 1990s are basically the only two periods of times that we see an obvious pattern of the decrease in income, and they all lasted for less than five years. This phenomenon is expected as during the time of early 1980s and 1990s, Canada was experiencing recession due to the change in monetary policy and a side effect of the cold war. As a result, the unemployment rate rose and due to an increase in the competitiveness of workforce, people's incomes have dropped. However, during other times, the income level have always demonstrated an increasing trend, indicating that without negative external forces, Canada generally has expanded its economy and enhanced its people's life quality well. The average after-tax income for economic families and unattached individuals in Canada has rose from around \$72400 in 1976 to \$96000 in 2020.

@@

Figure \@ref(fig:figure2) is an illustration of how the income level of the richest and poorest groups of economic families and unattached individuals differ between the year of 1976 and 2020. The blue line indicates the trend of the lowest decile of the average after-tax income in Canada, which barely increased. In fact, the lowest decile of average after-tax income was \$9200 in 1976, but only \$15000 after 44 years in 2020. On the other hand, the red line indicates the trend of the highest decile of the average after-tax income in Canada, which has a very clear increasing pattern, by looking at the data, it increased for $86000 between 1976 and 2020, which is more than 10 times the income level increase of the lowest decile. By comparing the two lines, we can see that the income inequality in Canada as a whole is getting larger as the two lines drift further away from each other as time passed.

@@

Prior to constructing the model, the dataset is split into the training set and the testing set with a proportion of 8:2. The training set is used to build the multiple linear regression models, and the testing set is used to test the accuracy and unbiasness of the model. R package "tidymodels"[@citetidymodels] is used to split the dataset.

@@

Two models are initially constructed based on the different possible factors that might have an effect on income inequality in Canada. The first model has continues variable year, categorical variable province as its predictor variables, and the difference between lowest and highest decile of income as its response variable. I would also like to check whether the variables year and province are implicitly related. Therefore, the second model has continues variable year, categorical variable province, and the interaction between them as its predictor variables, and the difference between lowest and highest decile of income as its response variable. To compare these two models to find out which is more accurate and is better at prediction, AIC and BIC tests are carried out, and $R^2$ is examined. Furthermore, the testing dataset is used to test how the two models perform when extra data is involved. RMSE is measured by the testing data to compare the prediction power of the two models. In Appendix \@ref(model-testing), these tests are performed and the test statistics of the two models are compared to select the model with better performance.

@@

  - $Y_{ij}$ is the difference between the highest and lowest decile of average after-tax income of economic families and unattched individuals of Canada in $i^{th}$ year and province $j$.
  - $\beta_{0}$ is the coefficient for intercept.
  - $\beta_{1}$ is the coefficient for the continuous year variable.
  - $\beta_{2}$ is the coefficient corresponding to province $j$.
  - $\beta_{3}$ is the coefficient for the interaction term between $i^{th}$ year and province $j$.
  - The baseline of this model is year 0 and Alberta province.

P-value is an essential value used to determine whether a predictor variable has significant impact on the response variable. It measures the probability that the observed data would occur given that the null hypothesis is true. In the context of a multiple linear regression model in this paper, null hypothesis is the hypothesis that there is no relationship between the predictor variable and the response variable [@cite6]. The common threshold of p-value is 0.05, so we will use this threshold in this paper. By examining the p-values of all the predictor variables of the model, we found that they are all less than 0.05, indicating that the hypothesis that there is no relationship between the variables of interest is rejected. As a result, none of the predictor variables of this model will be reduced because they all have some signifant impact on the response variable.

@@

Table \@ref(tab:table1) shows the coefficients of the predictor variables of Model \@ref(eq:1). Due to the fact that none of their corresponding p-values have a value more than 0.05, they are all kept in the final model. The baseline of the model is year 0 and Alberta province, and it also accounts for the interaction term between year and Alberta, which means that an intercept-only model presents the difference between lowest and highest decile of average after-tax income at year 0 in province Alberta. This is of course unrealistic because we only account for the data after 1976. Therefore, the "Year" variable has a strict minimum value of 1976 for this model. If the goal is to estimate the income inequality in the provinces other than Alberta, coefficient $\beta_2$ indicating which province it is and $\beta_3$ indicating the interaction between year and the specified province come into play, and their corresponding coefficient values can be found in Table \@ref(tab:table1) under "Coefficients".

@@

In section \@ref(results), we have talked about the outcome of the model, which is that time is the main factor expanding the issue of income disparity in Canada. Between 1976 and 2020, the difference in income level between the poorest group and the richest group has increase from \$184400 to \$264600 for approximately \$80200, which is even more than what an average Canadian earns in 2020. This implies that even though people are aware of this for quite some time, income disparity is still becoming a larger issue, and is likely to grow more in future. Some people claimed that the uneven distribution of income is inevitable if we want the country's economy to keep expanding; it's just a side effect of economic growth. It seems true in the case of Canada, since in Figure \@ref(fig:figure3), we can see that the only times when the difference between highest and lowest income decile decreased, is around 1980s and 1990s when Canada was experiencing recession. However, constant gdp growth does not mean that the average people of Canada are doing better, because commodity prices grow as well along with the income and may be a bigger pressure for the poorest people whose income had not increased much [@cite2]. The richest countries in the world typically share similar issues with respect to income inequality. On the contrary, countries that have good geographical locations and a potential to be more developed but choose to sacrifice that advantage for the well-being of the general public, such as lots of European countries, usually have pretty low income inequality and poverty level by their tax, welfare system and government policies.

As for Canada, it's just a matter of seeking a balance between economic development and equality. The federal income-tax system we currently have is progressive, which already favors the average people more than the rich [@cite8]. However, one of the solution to reduce income inequality more in federal level is to further increase the income-tax for the richest people. This could be the most effective approach, but even though it's enhancing the whole society's welfare, it's also harming the interest of some individuals, which could lead to further ethical issues and a whole lot judgments and discussions around this approach. Therefore, this solution would be too difficult and ideal to carry out.

@@

As we examined the issue of income inequality in different provinces earlier, we found that Alberta has the highest rate of increase in income inequality, followed by Newfoundland and Labrador and Ontario, while Quebec has the lowest increase in income inequality. These statistics in a way implies the economic development and opportunities in these provinces. For instance, Alberta and Ontario are the few provinces with the most GPD per capita and Market income per capita in recent years. Alberta has a whole  profitable oil and energy industry due to its rich oil resources, and Ontario, being the province with the largest population in Canada, developed its economy based on lots of sectors including manufacturing, exports, service, high-tech industries and more. In the provinces with higher income inequality, the residents are likely to experience more competition to seek for better opportunities, and are likely to be under higher level of stress, which could lead to more social and health problems. It's unrealistic to suspend the provinces' economic developments just to promote income equality. One approach to reduce it is to manipulate provincial income tax to favor the groups with lower incomes. Another possible solution is to encourage different provinces to carry out different policies to provide benefits and welfare to the lower-income groups to even out. These approaches could be challenging to carry out and unrealistic at the moment, but if actions are not taken, income disparity will likely to grow into more challenging problems in future.

@@

Strictly speaking, income level of individuals is counted as private information, and it's usually not allowed to publicly disclose it unless granted. Therefore, it would be hard to find individual's income information. The data used to justify income inequality in this paper is the difference between the lowest and highest decile of average after-tax income. This doesn't reflect the real gap between the richest and the poorest, because this dataset gathers the average of the top ten percent and the bottom ten percent. In reality, the issue of income inequality will only be worse that what's predicted by this paper. In the Canadian Income Survey, it indicates that less the people who live in remote areas which accounts for less than 2% of the population are not included in the target population of the survey. In the meanwhile, these people are usually the ones with lower income levels. As a result, a systematic error may occur with this dataset due to the survey's metholody and data collection process, which probably increase the lowest decile of average income and further weaken the issue of income inequality.

This paper analyzed how income diaparity in Canada is affected by time and different regions. However, income inequality is only an aspect of economic inequality. Other aspects of economic inequality includes uneven distribution of the wealth and assets of the Canadians, which also causes similar social challenges A possible future step of this paper is to focus on the wealth inequality and investigate whether it is a problem in Canada as well as its relationships with income inequality. In addition, in a few years from now, we could come back to this paper and see if the trend of income inequality is really what it predicted to be, or if the issue is lightened by potential government policies.

@@

AIC and BIC measure the prediction ability of the multiple linear regression model. AIC(Akaike’s Information Criteria) focuses on how well the model fits unknown data, while BIC(Bayesian Information Criteria) focuses on the true model and favours simpler models [@cite5]. Lower AIC and BIC both indicate that the model has better prediction power. By Table \@ref(tab:table2), Model 2 has slightly lower AIC and BIC than Model 1, implying that Model 2 has more prediction power.

@@

Check for assumption:
Assumptions to the multiple linear regression model are checked to ensure that the model is valid for this dataset. Figure \@ref(fig:figure5) shows the plots used to check for the assumption.

The Residuals vs Fitted plot checks for the linear relationship assumption. Since the red line is almost horizontal and there isn't any pattern, the model satisfies the linearity assumption. The Normal QQ plot checks for the residual normality assumption. Since almost all the dots are on the dashed line, the residuals follow a normal distribution. The Scale-Location plot checks for the homoscedasticity assumption. The red line is horizontal and the dots are evenly scattered, indicating that the variance of the residuals is constant [@cite4].

>>>>

In recent decades the world is developing fast economically, and one evidence of which is the rise in income level. As more countries become more developed in terms of their economy and eventually overcome their issues of poverty, their governments have gradually shift their goal to dealing with the problem of economic inequality. 
It concerns how the wealth and income of the population is distributed. Although we are generally living higher quality lives and getting higher incomes, another problem is being put forward by people who are claiming that the gap between the rich and poor is getting larger due to technological advancement, globalization, countries' governmental structure and policies, which betters off the people who are already wealthy by making it easy for them to seek for more opportunities, thus, snowballing their assets faster. Income disparity can be a huge issue. It makes the income per capita of a country less reflective of its people's true living quality, because income per capita does not tell the distribution of the income. The richest people may pull up the statistics about the average/median income level, but it doesn't mean the country is performing well economically as a whole. In addition to that, income inequality also has other potential negative effects. It could reduce the stability of the economy and increase the risk of financial crisis. It could also result in corruption, misallocation of resources as the rich is empowered economically, socially and even politically [@cite1].
Canada is one of the countries in the world that provides the most benefits and welfare, yet people still claimed that the economic inequality is rising along with the income over the past 2 decades [@cite2]. Therefore, I'm interested in whether income inequality is a serious issue in Canada. If so, in which regions and what period is it a serious issue? And is it possible to predict how the disparity of income change in future?

The data used in this report is obtained from Open Government Data of Statistics Canada. The dataset is called "Upper income limit, income share and average income by economic family type and income decile", and it provides different kinds of information about the income level of the households of Canada between 1976 and 2020 by the Canadian Income Survey. In this dataset, I'm interested in the year the data is from, the geographical location, the income decile so that we could observe the difference between the lowest and highest decile, and the actual value of the income. In section \@ref(data), a cleaned dataset is obtained and explained to perform further analysis. Visualizations in the forms of tables and figures are presented to help to explain the possible correlation between year, provinces and the income disparity in Canada. By data visualization, I've primarily assumed that the income disparity in Canada has generally increased as time passed and presents different trends in different provinces before constructing the model. In section \@ref(model), a multiple linear regression model is constructed to justify the relationship between year, province and income inequality, and to make future predictions. The interpretation of the final model along with all the findings regarding the evolution of income inequality is presented in section \@ref(results). A discussion is carried out in section \@ref(discussion) on the implications of the findings regarding time and geographical location, and possible solutions to this issue in federal and provincial level, as well as the weaknesses and future steps of this paper.

@@

The dataset, which is a summary of the income-related statistics in Canada between 1976 and 2020, is obtained from the Open Government Portal of Statistics Canada. The source of the data is from the Canadian Income Survey conducted annually. The survey covers all the possible population in Canada except for the people and households living in remote areas or indigenous settlements. It doesn't really affect the outcome of the survey, because these people only take up less than 2 percent of the population [@cite3]. The selected respondents were drawn from the Labour Force Survey samples, which is based on a stratified probability sampling [@cite3]. To reduce some non-sampling errors, telephone interviews were conducted prior to the main survey to increase the response rate of the selected participants Once they give consent to completing the survey, they will be able to complete it in an online form. Respondents are protected from the confidentiality rules, so that their privacy won't be disclosed after the survey is made public.
One possible bias came with the survey methodology was the over-coverage of the units that are not the target population and the under-coverage of certain sub-population that should be included in the survey, such as certain remote areas.

@@

In order to get further familiarized with the dataset and estimate the possible associations between the nation's average income and other factors, exploratory analysis is carried out by conducting data visualizations to observe whether the patterns between certain factors matches the generally expectations of the trend of income and income inequality in Canada. First, we will have a look through the overall trend of income level in Canada.

@@

As Figure \@ref(fig:figure1) has demonstrated, Canadians have experienced a rapidly growing income level since 1976 to now. Early 1980s and 1990s are basically the only two periods of times that we see an obvious pattern of the decrease in income, and they all lasted for less than five years. This phenomenon is expected as during the time of early 1980s and 1990s, Canada was experiencing recession due to the change in monetary policy and a side effect of the cold war. As a result, the unemployment rate rose and due to an increase in the competitiveness of workforce, people's incomes have dropped. However, during other times, the income level has always demonstrated an increasing trend, indicating that without negative external forces, Canada generally has expanded its economy and enhanced its people's life quality well. The average after-tax income for economic families and unattached individuals in Canada has rose from around \$72400 in 1976 to \$96000 in 2020.

@@

Figure \@ref(fig:figure2) is an illustration of how the income level of the richest and poorest groups of economic families and unattached individuals differ between the year of 1976 and 2020. The blue line indicates the trend of the lowest decile of the average after-tax income in Canada, which barely increased. In fact, the lowest decile of average after-tax income was \$9200 in 1976, but only \$15000 after 44 years in 2020. On the other hand, the red line indicates the trend of the highest decile of the average after-tax income in Canada, which has a very clear increasing pattern, by looking at the data, it increased for $86000 between 1976 and 2020, which is more than 10 times the income level increase of the lowest decile. By comparing the two lines, we can see that the income inequality in Canada is getting larger as the two lines drift further away from each other as time passed.

@@

Prior to constructing the model, the dataset is split into the training set and the testing set with a proportion of 8:2. The training set is used to build the multiple linear regression models, and the testing set is used to test the accuracy and unbiasedness of the model. R package "tidymodels"[@citetidymodels] is used to split the dataset.

@@

Two models are initially constructed based on the different possible factors that might have an effect on income inequality in Canada. The first model has continuous variable year, categorical variable province as its predictor variables, and the difference between lowest and highest decile of income as its response variable. I would also like to check whether the variables year and province are implicitly related. Therefore, the second model has continuous variable year, categorical variable province, and the interaction between them as its predictor variables, and the difference between lowest and highest decile of income as its response variable. To compare these two models to find out which is more accurate and is better at prediction, AIC and BIC tests are carried out, and $R^2$ is examined. Furthermore, the testing dataset is used to test how the two models perform when extra data is involved. RMSE is measured by the testing data to compare the prediction power of the two models. In Appendix \@ref(model-testing), these tests are performed and the test statistics of the two models are compared to select the model with better performance.

@@

  - $Y_{ij}$ is the difference between the highest and lowest decile of average after-tax income of economic families and unattached individuals of Canada in $i^{th}$ year and province $j$.
  - $\beta_{0}$ is the coefficient for intercept.
  - $\beta_{1}$ is the coefficient for the continuous year variable.
  - $\beta_{2}$ is the coefficient corresponding to province $j$.
  - $\beta_{3}$ is the coefficient for the interaction term between $i^{th}$ year and province $j$.
  - The baseline of this model is year 0 and Alberta province.

P-value is an essential value used to determine whether a predictor variable has significant impact on the response variable. It measures the probability that the observed data would occur given that the null hypothesis is true. In the context of a multiple linear regression model in this paper, null hypothesis is the hypothesis that there is no relationship between the predictor variable and the response variable [@cite6]. The common threshold of p-value is 0.05, so we will use this threshold in this paper. By examining the p-values of all the predictor variables of the model, we found that they are all less than 0.05, indicating that the hypothesis that there is no relationship between the variables of interest is rejected. As a result, none of the predictor variables of this model will be reduced because they all have some significant impact on the response variable.

@@

Table \@ref(tab:table1) shows the coefficients of the predictor variables of Model \@ref(eq:1). Since none of their corresponding p-values have a value more than 0.05, they are all kept in the final model. The baseline of the model is year 0 and Alberta province, and it also accounts for the interaction term between year and Alberta, which means that an intercept-only model presents the difference between lowest and highest decile of average after-tax income at year 0 in province Alberta. This is of course unrealistic because we only account for the data after 1976. Therefore, the "Year" variable has a strict minimum value of 1976 for this model. If the goal is to estimate the income inequality in the provinces other than Alberta, coefficient $\beta_2$ indicating which province it is and $\beta_3$ indicating the interaction between year and the specified province come into play, and their corresponding coefficient values can be found in Table \@ref(tab:table1) under "Coefficients".

@@

In section \@ref(results), we have talked about the outcome of the model, which is that time is the main factor expanding the issue of income disparity in Canada. Between 1976 and 2020, the difference in income level between the poorest group and the richest group has increase from \$184400 to \$264600 for approximately \$80200, which is even more than what an average Canadian earns in 2020. This implies that even though people are aware of this for quite some time, income disparity is still becoming a larger issue, and is likely to grow more in future. Some people claimed that the uneven distribution of income is inevitable if we want the country's economy to keep expanding; it's just a side effect of economic growth. It seems true in the case of Canada, since in Figure \@ref(fig:figure3), we can see that the only times when the difference between highest and lowest income decile decreased, is around 1980s and 1990s when Canada was experiencing recession. However, constant GDP growth does not mean that the average people of Canada are doing better, because commodity prices grow as well along with the income and may be a bigger pressure for the poorest people whose income had not increased much [@cite2]. The richest countries in the world typically share similar issues with respect to income inequality. On the contrary, countries that have good geographical locations and a potential to be more developed but choose to sacrifice that advantage for the well-being of the general public, such as lots of European countries, usually have pretty low income inequality and poverty level by their tax, welfare system and government policies.

As for Canada, it's just a matter of seeking a balance between economic development and equality. The federal income-tax system we currently have is progressive, which already favors the average people more than the rich [@cite8]. However, one of the solutions to reduce income inequality more in federal level is to further increase the income-tax for the richest people. This could be the most effective approach, but even though it's enhancing the whole society's welfare, it's also harming the interest of some individuals, which could lead to further ethical issues and a whole lot of judgments and discussions around this approach. Therefore, this solution would be too difficult and ideal to carry out.

@@

As we examined the issue of income inequality in different provinces earlier, we found that Alberta has the highest rate of increase in income inequality, followed by Newfoundland and Labrador and Ontario, while Quebec has the lowest increase in income inequality. These statistics in a way implies the economic development and opportunities in these provinces. For instance, Alberta and Ontario are the few provinces with the most GPD per capita and Market income per capita in recent years. Alberta has a whole profitable oil and energy industry due to its rich oil resources, and Ontario, being the province with the largest population in Canada, developed its economy based on lots of sectors including manufacturing, exports, service, high-tech industries and more. In the provinces with higher income inequality, the residents are likely to experience more competition to seek for better opportunities, and are likely to be under higher level of stress, which could lead to more social and health problems. It's unrealistic to suspend the provinces' economic developments just to promote income equality. One approach to reduce it is to manipulate provincial income tax to favor the groups with lower incomes. Another possible solution is to encourage different provinces to carry out different policies to provide benefits and welfare to the lower-income groups to even out. These approaches could be challenging to carry out and unrealistic at the moment, but if actions are not taken, income disparity will likely grow into more challenging problems in future.

@@

Strictly speaking, income level of individuals is counted as private information, and it's usually not allowed to publicly disclose it unless granted. Therefore, it would be hard to find individual's income information. The data used to justify income inequality in this paper is the difference between the lowest and highest decile of average after-tax income. This doesn't reflect the real gap between the richest and the poorest, because this dataset gathers the average of the top ten percent and the bottom ten percent. In reality, the issue of income inequality will only be worse that what's predicted by this paper. In the Canadian Income Survey, it indicates that less the people who live in remote areas which accounts for less than 2% of the population are not included in the target population of the survey. In the meanwhile, these people are usually the ones with lower income levels. As a result, a systematic error may occur with this dataset due to the survey's methodology and data collection process, which probably increase the lowest decile of average income and further weaken the issue of income inequality.

This paper analyzed how income disparity in Canada is affected by time and different regions. However, income inequality is only an aspect of economic inequality. Other aspects of economic inequality include uneven distribution of the wealth and assets of the Canadians, which also causes similar social challenges A possible future step of this paper is to focus on the wealth inequality and investigate whether it is a problem in Canada as well as its relationships with income inequality. In addition, in a few years from now, we could come back to this paper and see if the trend of income inequality is really what it predicted to be, or if the issue is lightened by potential government policies.

@@

AIC and BIC measure the prediction ability of the multiple linear regression model. AIC (Akaike’s Information Criteria) focuses on how well the model fits unknown data, while BIC (Bayesian Information Criteria) focuses on the true model and favours simpler models [@cite5]. Lower AIC and BIC both indicate that the model has better prediction power. By Table \@ref(tab:table2), Model 2 has slightly lower AIC and BIC than Model 1, implying that Model 2 has more prediction power.

@@

Check for assumptions:
Assumptions to the multiple linear regression model are checked to ensure that the model is valid for this dataset. Figure \@ref(fig:figure5) shows the plots used to check for the assumption.

The Residuals vs Fitted plot checks for the linear relationship assumption. Since the red line is almost horizontal and there isn't any pattern, the model satisfies the linearity assumption. The Normal QQ plot checks for the residual normality assumption. Since almost all the dots are on the dashed line, the residuals follow a normal distribution. The Scale-Location plot checks for the homoscedasticity assumption. The red line is horizontal, and the dots are evenly scattered, indicating that the variance of the residuals is constant [@cite4].
